### 2014-04-10 stat246/hw1.tex
s1

%To obtain the first term, note that it is a cross moment and can be obtained by applying $\frac{\partial^4}{\partial x_1^2\partial x_2^2}$ to the mgf of the distribution and evaluating at zero. The mgf itself is
%\[\exp\left(\frac{1}{2}\openm x_1&x_2\closem\openm v_1&a\\a&v_2\closem\openm x_1\\x_2\closem\right)=\exp\left(\frac{1}{2}(x_1^2v_1+2x_1x_2a+x_2^2v_2)\right)\]
%
%Differentiating this wrt $x_1$ twice and wrt $x_2$ once gives 
%\[ax_1v_1T+x_1v_1(x_1v_1+x_2a)(x_2v_2+x_1a)T+v_1(x_2v_2+x_1a)T\]
%where $T$ is the exponential term from above. Then, taking the derivatives and then evaluating gives $v_1v_2$, so the variance of $\h{a}$ is $\frac{1}{n}(v_1v_2-a^2)$.
Write $X_1^2X_2^2=X_1^2(\alpha X_1+Z)^2$. Then, the expectation of this is
\[\alpha^2E(X_1^4)+2\alpha E(Z)E(X_1^3)+E(X_1^2)E(Z^2)=\alpha^2E(X_1^4)+E(X_1^2)E(Z^2)\]
By differentiating the mgf for $X_1$ four times and evaluating at zero, we obtain $E(X_1^4)=3v_1$. We have that $E(X_1^2)=v_1$ and $E(Z^2)=\var(Z)$, so the resulting expression for $E(X_1^2X_2^2)$ is
\[3\frac{a^2}{v_1^2}v_1^2+v_1\left(v_2-\frac{a^2}{v_1}\right)=v_1v_2+2a^2\]
which gives us $\var(\h{a})=\frac{1}{n}(v_1v_2+a^2)$
\subsection*{6}
\ssn{a}
$\Sigma$ is positive definite because $A$ is nonsingular -- computing $x^T\Sigma x=x^TA^TAx=\|Ax\|_2^2$ will result in a positive value as long as $x$ is nonzero. The eigenvalues of $\Sigma$ are the squared eigenvalues of $A$, so the lowest eigenvalue of $\Sigma$ is $0.01$.
\ssn{b}
Sampling from a multivariate normal distribution with some specified covariance matrix can be done by first obtaining an iid sample of $n$ standard normal random variables, then transforming it via left-multiplication by $A^T$. Python code is attached.
\end{document}

>>>>

%To obtain the first term, note that it is a cross moment and can be obtained by applying $\frac{\partial^4}{\partial x_1^2\partial x_2^2}$ to the mgf of the distribution and evaluating at zero. The mgf itself is
%\[\exp\left(\frac{1}{2}\openm x_1&x_2\closem\openm v_1&a\\a&v_2\closem\openm x_1\\x_2\closem\right)=\exp\left(\frac{1}{2}(x_1^2v_1+2x_1x_2a+x_2^2v_2)\right)\]
%
%Differentiating this wrt $x_1$ twice and wrt $x_2$ once gives 
%\[ax_1v_1T+x_1v_1(x_1v_1+x_2a)(x_2v_2+x_1a)T+v_1(x_2v_2+x_1a)T\]
%where $T$ is the exponential term from above. Then, taking the derivatives and then evaluating gives $v_1v_2$, so the variance of $\h{a}$ is $\frac{1}{n}(v_1v_2-a^2)$.
Write $X_1^2X_2^2=X_1^2(\alpha X_1+Z)^2$. Then, the expectation of this is
\[\alpha^2E(X_1^4)+2\alpha E(Z)E(X_1^3)+E(X_1^2)E(Z^2)=\alpha^2E(X_1^4)+E(X_1^2)E(Z^2)\]
By differentiating the mgf for $X_1$ four times and evaluating at zero, we obtain $E(X_1^4)=3v_1$. We have that $E(X_1^2)=v_1$ and $E(Z^2)=\var(Z)$, so the resulting expression for $E(X_1^2X_2^2)$ is
\[3\frac{a^2}{v_1^2}v_1^2+v_1\left(v_2-\frac{a^2}{v_1}\right)=v_1v_2+2a^2\]
which gives us $\var(\h{a})=\frac{1}{n}(v_1v_2+a^2)$
\subsection*{6}
\ssn{a}
$\Sigma$ is positive definite because $A$ is nonsingular -- computing $x^T\Sigma x=x^TA^TAx=\|Ax\|_2^2$ will result in a positive value as long as $x$ is nonzero. The eigenvalues of $\Sigma$ are the squared eigenvalues of $A$, so the lowest eigenvalue of $\Sigma$ is $0.01$.
\ssn{b}
Sampling from a multivariate normal distribution with some specified covariance matrix can be done by first obtaining an iid sample of $n$ standard normal random variables, then transforming it via left-multiplication by $A^T$. Python code is attached.
\ssn{c,d}
Histograms for $e$: $d=100$ on the right, $d=10$ on the left. $n=500$ on top, $n=5000$ on bottom.

\includegraphics[width=0.5\textwidth]{hw1_files/e10-500.png}
\includegraphics[width=0.5\textwidth]{hw1_files/e100-500.png}

\includegraphics[width=0.5\textwidth]{hw1_files/e10-5000.png}
\includegraphics[width=0.5\textwidth]{hw1_files/e100-5000.png}

The mean and stddev for $e$ are in the following table (presented as a pair mean,stddev):

\begin{tabular}{c|c|c}
    &$d=10$&$d=100$\\
    \hline
    $n=500$&$0.447,64.85$&$0.0302,60.67$\\
    \hline
    $n=5000$&$-4.137,204.75$&$0.00706,200.00$\\
\end{tabular}

Histograms for the eigenvalue estimates: $d=100$ on right, $d=10$ on left

\includegraphics[width=0.5\textwidth]{hw1_files/m10-500.png}
\includegraphics[width=0.5\textwidth]{hw1_files/m100-500.png}

\includegraphics[width=0.5\textwidth]{hw1_files/m10-5000.png}
\includegraphics[width=0.5\textwidth]{hw1_files/m100-5000.png}

The table of means and standard deviations:

\begin{tabular}{c|c|c}
    &$d=10$&$d=100$\\
    \hline
    $n=500$&$-0.0214,0.0626$&$-0.274,0.0669$\\
    \hline
    $n=5000$&$-0.00140,0.0205$&$-0.0239,0.0197$\\
\end{tabular}

As expected, the standardized errors cluster around zero with an apparently normal distribution. The standard deviation is lower for $d=100$ than $d=10$ and higher for $n=5000$ than $n=500$. The eigenvalue ratios indicate that the highest eigenvalue of the inverse of the sample covariance is consistently lower than that of the highest eigenvalue of the actual.
\end{document}
